---
title: "BGC_class_barplot"
author: "Minghao Chia"
date: "12/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries

```{r}
library(tidyverse)
library(ComplexUpset)
```

Define helper function

```{r}

fix_labels2 <- function(df){
df$TYPE <- ifelse(df$TYPE %in% c("lanthipeptide","bacteriocin", "thiopeptide","lassopeptide","cyanobactin","sactipeptide",
"linaridin","bottromycin","microviridin","head_to_tail","glycocin",
"LAP","lipolanthine","proteusin","microcin", "RaS-RiPP"), "RiPP", df$TYPE )
df$TYPE <- ifelse(df$TYPE == "NRP", "NRPS", df$TYPE )
df$TYPE <- ifelse(df$TYPE %in% c("T1PKS","T3PKS", "transAT-PKS", "hglE-KS"), 
                                  "Polyketide", df$TYPE)
df$TYPE <- ifelse(df$TYPE %in% c("oligosaccharide", "amglyccycl"), "Saccharide", df$TYPE )
df$TYPE <- ifelse(df$TYPE == "terpene", "Terpene", df$TYPE )
df$TYPE <- ifelse(df$TYPE %in% c("arylpolyene","betalactone","TfuA-related","butyrolactone",
                                                           "ladderane","CDPS", "siderophore", "phenazine", "resorcinol",
                                                           "phosphonate"), "Other", df$TYPE )

df$TYPE <- ifelse(str_detect(string=df$TYPE,pattern="[;-]"), "Hybrid", df$TYPE)

return(df)
}

```


Code to generate fig 2D, Fig S13 and Fig S14. 

Load metadata and input data for figure making

```{r, message=FALSE}

bgc_novel_pdt_count <- read_tsv(file = "../tables/fig2D_top_bgc_novel_pdf_count.tsv", show_col_types = FALSE)

GCF_novel_pdt_count <- read_tsv(file = "../tables/fig2D_bot_GCF_novel_pdf_count.tsv", show_col_types = FALSE)

all_GCF_uniq_fmt <- read_tsv(file="../tables/all_GCF_uniq_fmt.tsv", show_col_types = FALSE)

final_rep_clusters <- read_tsv(file="../tables/final_rep_clusters.tsv", show_col_types = FALSE )

```


Stacked barcharts showing the number of BGCs (Fig 2D top) and GCFs (Fig 2D bottom) in different product classes that are known or novel in SPMP. Piecharts show the overall breakdown.

Plot simplified product class distribution for novel and non-novel BGCs

Figure 2D top

```{r message=FALSE}

bgc_novel_pdt_barplot <- ggplot(bgc_novel_pdt_count, 
                                      aes(x=reorder(as.factor(TYPE), -freq), 
                                          y=freq, fill=novel)) + 
  geom_col(width=0.3) + theme_classic() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=12), axis.text.y = element_text(size = 12)) +
scale_x_discrete(limits=c("Saccharide", "RiPP", "Polyketide", "NRPS", "Terpene", "Hybrid", "Other" ))  +
theme(axis.title=element_text(size=12)) + ylab("# BGC") + xlab(NULL) + theme(aspect.ratio = 1/1.3) +
  scale_fill_manual(values= c("#E6A66D", "#6F9AB8"))

bgc_novel_pdt_barplot

#Pi chart
#https://ggplot2.tidyverse.org/reference/coord_polar.html

bgc_novel_pie <-  ggplot(plyr::count(all_GCF_uniq_fmt,"novel"), 
                         aes(x="", y=freq, fill=novel)) + geom_col(width=1) + 
  coord_polar("y", start= pi/3) +  scale_fill_manual(values= c("#E6A66D","#6F9AB8"))

bgc_novel_pie


#Fractions in the pi chart

bgc_novel_stats <- plyr::count(all_GCF_uniq_fmt,"novel")

sum(bgc_novel_stats$freq) #27084

bgc_novel_stats %>% filter(novel == TRUE) %>% pull(freq) %>% sum(.)/sum(bgc_novel_stats$freq) #0.882


```

Figure 2D bottom

```{r echo=FALSE}

GCF_novel_pdt_barplot <- ggplot(GCF_novel_pdt_count, aes(x=reorder(as.factor(TYPE), -freq), 
                                                                     y=freq, fill=novel)) + 
  geom_col(width=0.3) + theme_classic() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=12), axis.text.y = element_text(size = 12)) +
scale_x_discrete(limits=c("Saccharide", "RiPP", "Polyketide", "NRPS", "Terpene", "Hybrid", "Other" ))  +
  scale_y_continuous(breaks=c(0,1000,2000,3000,4000,5000,6000)) +
theme(axis.title=element_text(size=12)) + 
  ylab("# GCF") + xlab(NULL) + theme(aspect.ratio = 1/1.3) +
  scale_fill_manual(values= c("#E6A66D", "#6F9AB8"))
 
GCF_novel_pdt_barplot # bottom panel for corrections

#piechart for panel 2D bottom

GCF_novel_pie <-  ggplot(plyr::count(final_rep_clusters,"novel"), 
                         aes(x="", y=freq, fill=novel)) + geom_col(width=1) + 
  coord_polar("y", start= pi/3) +  scale_fill_manual(values= c("#E6A66D", "#6F9AB8"))

GCF_novel_pie 

#Fractions in the pi chart

GCF_novel_stats <- plyr::count(final_rep_clusters,"novel")

sum(GCF_novel_stats$freq) #16055

GCF_novel_stats %>% filter(novel == TRUE) %>% pull(freq) %>% sum(.)/sum(GCF_novel_stats$freq) #0.9426

```

Figure S13.
```{r}

ASbgc_novel_pdt_count <- plyr::count(fix_labels2(all_GCF_uniq_fmt %>% dplyr::filter(detector == "antismash")), 
                                           c("TYPE", "novel"))

ASbgc_novel_pdt_barplot <- ggplot(ASbgc_novel_pdt_count, 
                                      aes(x=reorder(as.factor(TYPE), -freq), 
                                          y=freq, fill=novel)) + 
  geom_col(width=0.3) + theme_classic() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=12), axis.text.y = element_text(size = 12)) +
scale_x_discrete(limits=c("Saccharide", "RiPP", "Polyketide", "NRPS", "Terpene", "Hybrid", "Other" ))  +
theme(axis.title=element_text(size=12)) + ylab("# BGC") + xlab(NULL) + theme(aspect.ratio = 1/1.3) +
  scale_fill_manual(values= c("#E6A66D","#6F9AB8"))

ASbgc_novel_pdt_barplot


# Pi chart


ASbgc_novel_pie <-  ggplot(plyr::count(all_GCF_uniq_fmt %>% dplyr::filter(detector == "antismash"),"novel"), 
                         aes(x="", y=freq, fill=novel)) + geom_col(width=1) + 
  coord_polar("y", start= pi/3) +  scale_fill_manual(values= c("#E6A66D","#6F9AB8"))

ASbgc_novel_pie

#Fractions in the pi chart

ASbgc_novel_stats <- plyr::count(all_GCF_uniq_fmt %>% dplyr::filter(detector == "antismash"),"novel")

sum(ASbgc_novel_stats$freq) #3909

ASbgc_novel_stats %>% filter(novel == TRUE) %>% pull(freq) %>% sum(.)/sum(ASbgc_novel_stats$freq) #0.36377


```
Fig S14

```{r}

categories <- c("in_curated_db", "in_HRGM", "antismash_SPMP_GCF", "novel")
names(categories) <- categories

GCF_upset_df <- final_rep_clusters %>% dplyr::filter(detector=="antismash") %>% dplyr::select(detector, in_curated_db, in_HRGM, novel)
GCF_upset_df$antismash_SPMP_GCF <- ifelse(GCF_upset_df$detector=="antismash", TRUE, FALSE)

GCF_upset_df <- GCF_upset_df %>% dplyr::select(-detector)

#https://stackoverflow.com/questions/66323268/problem-with-upset-plot-intersection-numbers
#https://krassowski.github.io/complex-upset/articles/Examples_R.html
GCF_upset_plot <- upset(data=GCF_upset_df, intersect=categories,
      width_ratio = 0.125, mode ="intersect",
      sort_intersections = FALSE,
      intersections = list( "antismash_SPMP_GCF", 
                           c("antismash_SPMP_GCF", "novel"),
                           c("antismash_SPMP_GCF", "in_HRGM"),
                           c("antismash_SPMP_GCF", "in_curated_db"),
                           c("antismash_SPMP_GCF", "in_curated_db", "in_HRGM")),
      stripes="white")

GCF_upset_plot

#Pi chart for fig S14
ASgcf_novel_pie <-  ggplot(plyr::count(final_rep_clusters %>% dplyr::filter(detector == "antismash"),"novel"), 
                         aes(x="", y=freq, fill=novel)) + geom_col(width=1) + 
  coord_polar("y", start= pi/3) +  scale_fill_manual(values= c("#E6A66D", "#6F9AB8"))

ASgcf_novel_pie


#Fractions in the pi chart

ASgcf_novel_stats <- plyr::count(final_rep_clusters %>% dplyr::filter(detector == "antismash"),"novel")

sum(ASgcf_novel_stats$freq) #1010

ASgcf_novel_stats %>% filter(novel == TRUE) %>% pull(freq) %>% sum(.)/sum(ASgcf_novel_stats$freq) #0.487



```


